Learning \strips\ action models from {\em state-invariants}

### 2019-02-06 doc/ijcai-19-deductivelearning/deductivelearning.tex
complete version for sections 1,2 and 3 ijcai19 paper on deductive learning

>>>>

\subsection{Improving the compilation for {\em domain mutex}}
We define a {\em domain mutex} as $p\leftrightarrow \neg q$ universally quantified formula where $p\in\Psi$ and $q\in\Psi$ are predicates that shape the fluents of a given planning problem. For instance, $\forall x_1\ holding(x_1)\leftrightarrow\neg clear(x_1)$  or $\forall x_1\ holding(x_1)\leftrightarrow\neg ontable(x_1)$ are {\em domain mutex} of the {\em blocksworld} domain.

We pay special attention to this particular class of {\em state-invariants} because they are associated to the different {\em state-properties} of a given object and because they can be effectively exploited to improve our compilation for the learning /strips/ action models. Our approach is to extend the conditional effects of the $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertPre_{p,\xi}}$ actions that determine the learned model with extra conditional effects indicating that the programmed model is {\em invalid} if it is inconsistent with a given {\em domain mutex}. Note that this {\em consistency} checking is more effective than the one implemented at the $\mathsf{apply_{\xi,\omega}}$ since it happens at an earlier stage of the planning process.

Formally, given a $p\leftrightarrow \neg q$ we extend the actions for setting a precondition $p$ in a given action schema $\xi$ as follows:

\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p}(\xi), \neg eff_{p}(\xi),\\
&mode_{insert},\neg mode_{inval}\},\\
\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre_{p}(\xi)\},\\
&\{pre_{q}(\xi)\}\rhd\{mode_{inval}\}.
\end{align*}
\end{small}
The same procedure is applied for action $insertPre_{q,\xi}$ to ban this programming action iff $pre_{q}(\xi)$ precondition is already set. A similar procedure is also applied to $\mathsf{insertEff_{p,\xi}}$ and $\mathsf{insertEff_{q,\xi}}$ actions for banning two {\em negative effects} or (two {\em positive effects}) that are considered {\em domain mutex}.
